---
title: "Landuse Southern Germany"
author: "Lucia Tischer"
date: "18 April 2018"
output: ioslides_presentation
fontsize: 8pt
---




```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=5.5, fig.align='center')
```

```{r, eval=TRUE}
setwd("C:/Users/Ltischer/Documents/Studium/A Master/Geostatistics/Deforestation_SouthernGermany")
library(raster)
library(RStoolbox)
library(rgdal)
library(ggplot2)
library(gridExtra)
library(rgeos)
library(ggpubr)
library(car)
```

***

### 1. Question and Chosen Data

**How did the Landscape around my Hometown change during the last 30 Years?**
  
  - Data Download from: https://earthexplorer.usgs.gov/
  - Landsat 4/5 TM, Level 2
  - Cloudfree Images available for 20.07.1988 and 04.07.2011 (23 years difference)
  - Scene of path 193, row 27 cover the area around my Hometown (Marktoberdorf)
    
    - Landcover in 1988 and 2011?
    - Productivity of forest and agricultural Land in 1988 and 2011?
    - Differences of productivity within different elevation classes?


***

### 2. Expectations

  + Transformation of Landscape
    - less forest
    - less agriculture
    - more urban areas
  
  + Differences of productiviy
    - ~ landcover: prod(forest)  >  prod(agriculture)
    - ~ year: prod(1988)  >  prod(2011)
    - ~ elevation: prod(low)  >  prod(high)

***

### 3. Where is Marktoberdorf?
```{r, eval=TRUE}
# download elevation data and borders of germany
germany_border <- getData("GADM", country = "DEU", level = 1)
elev_germany <- getData('alt', country = "DEU") # download the elvation data

# define the coordinates of Marktoberdorf and W端rzburg
gps_Marktoberdorf <- as.data.frame(cbind( x=10.6158265, y=47.778483, id=1))
coordinates(gps_Marktoberdorf) <- c("x", "y")
Marktoberdorf <- as.data.frame(cbind( x=12.7, y=47.778483, id=1))
coordinates(Marktoberdorf) <- c("x", "y")

gps_Wurzburg <- as.data.frame(cbind( x=9.9533548, y=49.7913044, id=1))
coordinates(gps_Wurzburg) <- c("x", "y")
Wurzburg <- as.data.frame(cbind( x=11.4, y=49.7913044, id=1))
coordinates(Wurzburg) <- c("x", "y")

# define the projection of the coordinates
proj4string(gps_Marktoberdorf) <- CRS("+proj=longlat +datum=WGS84")
gps_Marktoberdorf <- spTransform(gps_Marktoberdorf, CRS(projection(elev_germany)))
proj4string(gps_Wurzburg) <- CRS("+proj=longlat +datum=WGS84")
gps_Wurzburg <- spTransform(gps_Wurzburg, CRS(projection(elev_germany)))



plot(elev_germany, 
     main="Germany with Marktoberdorf and W端rzburg",
     legend=FALSE,  
     las=1, 
     xaxt="n", yaxt="n")
plot(germany_border, 
     main="Germany with Marktoberdorf and W端rzburg",
     legend=FALSE, 
     add=TRUE, 
     las=1, 
     xaxt="n", yaxt="n")
plot(gps_Marktoberdorf, 
     add=TRUE, 
     size=5, 
     col="darkred", pch=16)
plot(gps_Wurzburg, 
     add=TRUE, 
     size=5, 
     col="darkred", pch=16)
text(x=Marktoberdorf, 
     labels="Marktoberdorf")
text(x=Wurzburg, 
     labels="W端rzburg")
```


```{r, eval=TRUE}
########## 0. partly data from 1988 (needed to set projection of gps-data etc.)
setwd("C:/Users/Ltischer/Documents/Studium/A Master/Geostatistics/Deforestation_SouthernGermany")
p193r27_1988_xml <- readMeta("raster/LT05_p193r27_1988_07_20/LT05_L1TP_193027_19880720_20170208_01_T1.xml")
p193r27_1988_cdr <- stackMeta(p193r27_1988_xml, quantity=c("sre", "bt")) # --> no bt-bands, only sre-bands are loaded



########## 1. Reading in and transforming the coordinates of Marktoberdorf
# Therefore: Read in the data from 1988 first for projection! (see below)

gps_Marktoberdorf <- as.data.frame(cbind( x=10.6158265, y=47.778483, id=1))

coordinates(gps_Marktoberdorf) <- c("x", "y")

#projection(gps_Marktoberdorf)
#projection(p193r27_1988_cdr)

proj4string(gps_Marktoberdorf) <- CRS("+proj=longlat +datum=WGS84")
gps_Marktoberdorf <- spTransform(gps_Marktoberdorf, CRS(projection(p193r27_1988_cdr)))

#gps_Marktoberdorf





########## 2. Creating a 25km-Buffer around Marktoberdorf
buf_25 <- gBuffer(gps_Marktoberdorf, width=25000, byid=TRUE) 

# prepare the polygons for plotting
buf_25_df <- fortify(buf_25, region="id") # transform the buffer-data to dataframe
buf_25_df_final <- merge(buf_25_df, buf_25@data, by="id") # add the information

# transform also coordinates back to dataframe for plotting
gps_Marktoberdorf_df <- as.data.frame(gps_Marktoberdorf) 

```

```{r}
plot(buf_25)
plot(gps_Marktoberdorf, add=TRUE)





########## 3. Data from 1988 
setwd("C:/Users/Ltischer/Documents/Studium/A Master/Geostatistics/Deforestation_SouthernGermany")
p193r27_1988_xml <- readMeta("raster/LT05_p193r27_1988_07_20/LT05_L1TP_193027_19880720_20170208_01_T1.xml")
p193r27_1988_cdr <- stackMeta(p193r27_1988_xml, quantity=c("sre", "bt")) # --> no bt-bands, only sre-bands are loaded

scaleF_1988 <- getMeta(p193r27_1988_cdr, p193r27_1988_xml, what="SCALE_FACTOR")
scaleF_1988

# transform DNs back to real values
p193r27_1988_cdr <- p193r27_1988_cdr*scaleF_1988

# plot the data from 1985 and create an extent to crop
ggRGB(p193r27_1988_cdr, stretch="lin")+
  ggtitle("1988")


# create the extent of the 25km-buffer and crop to this extent
ex <- extent(buf_25)
p193r27_1988_cdr_crop <- crop(p193r27_1988_cdr, extent(buf_25))



# save the results
writeRaster(p193r27_1988_cdr, "results/p193r27_1988_cdr.grd")
writeRaster(p193r27_1988_cdr_crop, "results/p193r27_1988_cdr_crop.grd")




########## 4. Data from 2011
p193r27_2011_xml <- readMeta("raster/LT05_p193r27_2011_07_04/LT05_L1TP_193027_20110704_20161008_01_T1.xml")
p193r27_2011_cdr <- stackMeta(p193r27_2011_xml, quantity=c("sre", "bt")) # --> no bt-bands, only sre-bands are loaded

scaleF_2011 <- getMeta(p193r27_2011_cdr, p193r27_2011_xml, what="SCALE_FACTOR")
scaleF_2011

# transform DNs back to real values
p193r27_2011_cdr <- p193r27_2011_cdr*scaleF_2011 

# plot all the scene from 2011
ggRGB(p193r27_2011_cdr, stretch="lin", r=3, g=2, b=1)+
  ggtitle("2011")


# create the extent of the 25 k buffer
p193r27_2011_cdr_crop <- crop(p193r27_2011_cdr, extent(buf_25))



# save the results
writeRaster(p193r27_2011_cdr, "results/p193r27_2011_cdr.grd")
writeRaster(p193r27_2011_cdr_crop, "results/p193r27_2011_cdr_crop.grd")

```


```{r}
# plot the Coordinates of Marktoberdorf and 25 km Buffer around
# therefore load the results calculated before
setwd("C:/Users/Ltischer/Documents/Studium/A Master/Geostatistics/Deforestation_SouthernGermany")
p193r27_2011_cdr <- brick("results/p193r27_2011_cdr.grd")
p193r27_2011_cdr_crop <- brick("results/p193r27_2011_cdr_crop.grd")

# plot the data from 1988 and create an extent to crop
g1a <- ggRGB(p193r27_2011_cdr,
      stretch="lin", geom_raster=TRUE) +
  ggtitle("2011 complete") +
  theme(plot.title=element_text(size=12, colour="black", face="bold"), 
        legend.title=element_text(size=10, colour = "black", face="bold")) +
  geom_polygon(data=buf_25_df_final, 
               mapping=aes(x=buf_25_df_final$long, 
                           y=buf_25_df_final$lat,
                           group=group), 
               colour="orange", 
               fill="orange", 
               alpha=0.3) +
  geom_point(data=gps_Marktoberdorf_df,
             aes(x=gps_Marktoberdorf_df$x,y=gps_Marktoberdorf_df$y), 
             colour="red", 
             shape=3, size=3, stroke=1.5)

# plot the data from 2011
g1b <- ggRGB(p193r27_2011_cdr_crop,
      stretch="lin", geom_raster=TRUE) +
  ggtitle("2011 croped") +
  theme(plot.title=element_text(size=12, colour="black", face="bold"), 
        legend.title=element_text(size=10, colour = "black", face="bold")) +
  geom_polygon(data=buf_25_df_final, 
               mapping=aes(x=buf_25_df_final$long, 
                           y=buf_25_df_final$lat,
                           group=group), 
               colour="orange", 
               fill="orange", 
               alpha=0.3) +
  geom_point(data=gps_Marktoberdorf_df,
             aes(x=gps_Marktoberdorf_df$x,y=gps_Marktoberdorf_df$y), 
             colour="red", 
             shape=3, size=3, stroke=1.5)

grid.arrange(g1a, g1b, ncol=2, nrow=1)

```


***


### 4. Preparing the Data 
  
  + Calculation of Real Values from DNs
  + Removal of Artefacts (values outside of [0,1])
  + Topografic Illumation Correction  
    - RStoolbox::topCor()
    - method: "C"
    - resampled elevation data from raster::getData()
  
  + no clouds removed (no clouds detected in chosen extracts)


***


### Artefacts?
```{r, eval=TRUE}
setwd("C:/Users/Ltischer/Documents/Studium/A Master/Geostatistics/Deforestation_SouthernGermany")

# load the results calculated earlier
p193r27_1988_cdr <- brick("results/p193r27_1988_cdr.grd")
p193r27_2011_cdr <- brick("results/p193r27_2011_cdr.grd")

par(mfrow=c(1,2), las=2)
########## 1. Data from 1988
boxplot(p193r27_1988_cdr, main="Values Bands 1-6, p193r27\n(1988)")
lines(x=c(0,8), y=c(1,1), col="red", lty="dashed")
lines(x=c(0,8), y=c(0,0), col="red", lty="dashed")
## --> some single data points are outside of the range [0,1]

########## 2. Data from 2011
boxplot(p193r27_2011_cdr, main="Values Bands 1-6, p193r27\n(2011)")
lines(x=c(0,8), y=c(1,1), col="red", lty="dashed")
lines(x=c(0,8), y=c(0,0), col="red", lty="dashed")
## --> some single data points are outside of the range [0,1]

```

```{r}
########## 1. Data from 1988
summary(p193r27_1988_cdr)
## --> some single data points are outside of the range [0,1]

# remove these points from the data
p193r27_1988_cdr_new <- p193r27_1988_cdr # create a copy of the data
query_raster_p193r27_1988 <- ((p193r27_1988_cdr_new[[1:6]] < 0) | (p193r27_1988_cdr_new[[1:6]] > 1)) # create a query raster, which defines the values outside of the Intervall [0, 1]
p193r27_1988_cdr_new[query_raster_p193r27_1988] <- NA # set all these values to zero

# plot the new data again
boxplot(p193r27_1988_cdr_new, 
        main="Values Bands 1-6, p193r27_new from 1988", 
        ylim=c(-0.2,1.2))
lines(x=c(0,8), y=c(1,1), col="red", lty="dashed")
lines(x=c(0,8), y=c(0,0), col="red", lty="dashed")


# create the extract of 25-km-buffer from the data and plot it
p193r27_1988_cdr_new_crop <- crop(p193r27_1988_cdr_new, extent(buf_25))
ggRGB(p193r27_1988_cdr_new_crop, stretch="lin") +
  ggtitle("Cropped Scene from 1988\n(25 km-buffer around Marktoberdorf)") +
  theme(plot.title=element_text(size=12, colour="black", face="bold"))

# create also a mask for only the buffer area
p193r27_1988_cdr_new_crop_mask <- mask(p193r27_1988_cdr_new_crop, buf_25)
ggRGB(p193r27_1988_cdr_new_crop_mask, stretch="lin") +
  ggtitle("Cropped Scene from 1988\n(ONLY 25 km-buffer around Marktoberdorf)") +
  theme(plot.title=element_text(size=12, colour="black", face="bold"))


# save the results
writeRaster(p193r27_1988_cdr_new, "results/p193r27_1988_cdr_new.grd")
writeRaster(p193r27_1988_cdr_new_crop, "results/p193r27_1988_cdr_new_crop.grd")
writeRaster(p193r27_1988_cdr_new_crop_mask, "results/p193r27_1988_cdr_new_crop_mask.grd")





########## 2. Data from 2011
summary(p193r27_2011_cdr)
## --> some single data points are outside of the range [0,1]

# remove these points from the data
p193r27_2011_cdr_new <- p193r27_2011_cdr # create a copy of the data
query_raster_p193r27_2011 <- ((p193r27_2011_cdr_new[[1:6]] < 0) | (p193r27_2011_cdr_new[[1:6]] > 1)) # create a query raster, which defines the values outside of the Intervall [0, 1]
p193r27_2011_cdr_new[query_raster_p193r27_2011] <- NA # set all these values to zero

# plot the new data again
boxplot(p193r27_2011_cdr_new, 
        main="Values Bands 1-6, p193r27_new from 2011", 
        ylim=c(-0.2, 1.2))
lines(x=c(0,8), y=c(1,1), col="red", lty="dashed")
lines(x=c(0,8), y=c(0,0), col="red", lty="dashed")


# create the extract of 25-km-buffer from the data and plot it
p193r27_2011_cdr_new_crop <- crop(p193r27_2011_cdr_new, extent(buf_25))
ggRGB(p193r27_2011_cdr_new_crop, stretch="lin") +
  ggtitle("Cropped Scene from 2011\n(25 km-buffer around Marktoberdorf)") +
  theme(plot.title=element_text(size=12, colour="black", face="bold"))

# create also a mask for only the buffer area
p193r27_2011_cdr_new_crop_mask <- mask(p193r27_2011_cdr_new_crop, buf_25)
ggRGB(p193r27_2011_cdr_new_crop_mask, stretch="lin") +
  ggtitle("Cropped Scene from 2011\n(ONLY 25 km-buffer around Marktoberdorf)") +
  theme(plot.title=element_text(size=12, colour="black", face="bold"))


# save the results
writeRaster(p193r27_2011_cdr_new, "results/p193r27_2011_cdr_new.grd")
writeRaster(p193r27_2011_cdr_new_crop, "results/p193r27_2011_cdr_new_crop.grd")
writeRaster(p193r27_2011_cdr_new_crop_mask, "results/p193r27_2011_cdr_new_crop_mask.grd")
```

***

### Topografic Illumination Correction?
```{r}

# load the scenes if calculated before
setwd("C:/Users/Ltischer/Documents/Studium/A Master/Geostatistics/Deforestation_SouthernGermany")
p193r27_1988_cdr_new_crop_mask <- brick("results/p193r27_1988_cdr_new_crop_mask.grd")
p193r27_2011_cdr_new_crop_mask <- brick("results/p193r27_2011_cdr_new_crop_mask.grd")

# load also data which is not masked yet (for adopting the resolution of the elevation data)
p193r27_1988_cdr_new_crop <- brick("results/p193r27_1988_cdr_new_crop.grd")
p193r27_2011_cdr_new_crop <- brick("results/p193r27_2011_cdr_new_crop.grd")


# download the elevation data for germany
x <- getData("ISO3") # store all the country codes as variable
x$ISO3[x$NAME == "Germany"] # query the country code for Germany
elev_germany <- getData('alt', country = "DEU") # download the elvation data

# compare the projections of elevation data and Marktoberdorf data
identical(projection(elev_germany), projection(p193r27_1988_cdr_new_crop_mask))
identical(projection(elev_germany), projection(p193r27_2011_cdr_new_crop_mask)) # --> the projections differ!

# reproject the elevation data
elev_germany_proj <- projectRaster(elev_germany, crs=crs(p193r27_1988_cdr_new_crop_mask))
identical(projection(elev_germany_proj), projection(p193r27_2011_cdr_new_crop_mask))
identical(projection(elev_germany_proj), projection(p193r27_1988_cdr_new_crop_mask)) # --> now they are identical

# crop the elevation data to scene-size
elev_germany_proj_crop <- crop(elev_germany_proj, p193r27_1988_cdr_new_crop_mask)
elev_germany_proj_crop_mask <- mask(elev_germany_proj_crop, buf_25)

# compare the resolution and dimension of elevation data and scene data
elev_germany_proj_crop
p193r27_1988_cdr_new_crop # --> elevation data has different dimensions and different resolution (non-square)

# resample the elevation data to adopt resolution and dimension to scene-data
elev_germany_proj_crop_resampled <- resample(elev_germany_proj_crop, p193r27_1988_cdr_new_crop, method="bilinear")

# check resolution and dimensions again
elev_germany_proj_crop_resampled
p193r27_1988_cdr_new_crop # --> parameters are identical now

# mask the new elevation data to shape of 25km-buffer
elev_germany_proj_crop_resampled_mask <- mask(elev_germany_proj_crop_resampled, buf_25)


# save the created elevations data
writeRaster(elev_germany, "results/elev_germany.grd")
writeRaster(elev_germany_proj_crop_mask, "results/elev_germany_proj_crop_mask.grd")
writeRaster(elev_germany_proj_crop_resampled, "results/elev_germany_proj_crop_resampled.grd")
writeRaster(elev_germany_proj_crop_resampled_mask, "results/elev_germany_proj_crop_resampled_mask.grd")


# load again the metadata for the two scenes
p193r27_1988_xml <- readMeta("raster/LT05_p193r27_1988_07_20/LT05_L1TP_193027_19880720_20170208_01_T1.xml")
p193r27_2011_xml <- readMeta("raster/LT05_p193r27_2011_07_04/LT05_L1TP_193027_20110704_20161008_01_T1.xml")


# apply the topografic illumination correction
p193r27_1988_cdr_new_crop_illu <- topCor(p193r27_1988_cdr_new_crop, 
                                         dem=elev_germany_proj_crop_resampled, 
                                         metaData=p193r27_1988_xml, 
                                         method="C")

p193r27_2011_cdr_new_crop_illu <- topCor(p193r27_2011_cdr_new_crop, 
                                         dem=elev_germany_proj_crop_resampled, 
                                         metaData=p193r27_2011_xml, 
                                         method="C")



# mask the corrected images to buffer-zone
p193r27_1988_cdr_new_crop_illu_mask <- mask(p193r27_1988_cdr_new_crop_illu, buf_25)
ggRGB(p193r27_1988_cdr_new_crop_illu_mask, stretch="lin") +
  ggtitle("1988, illumination corrected, 25-km-buffer around Marktoberdorf")

p193r27_2011_cdr_new_crop_illu_mask <- mask(p193r27_2011_cdr_new_crop_illu, buf_25)
ggRGB(p193r27_2011_cdr_new_crop_illu_mask, stretch="lin") +
  ggtitle("2011, illumination corrected, 25-km-buffer around Marktoberdorf")



# save the results
writeRaster(p193r27_1988_cdr_new_crop_illu, "results/p193r27_1988_cdr_new_crop_illu.grd")
writeRaster(p193r27_1988_cdr_new_crop_illu_mask, "results/p193r27_1988_cdr_new_crop_illu_mask.grd")
writeRaster(p193r27_2011_cdr_new_crop_illu, "results/p193r27_2011_cdr_new_crop_illu.grd")
writeRaster(p193r27_2011_cdr_new_crop_illu_mask, "results/p193r27_2011_cdr_new_crop_illu_mask.grd")
```

```{r, eval=TRUE}
# load the results calculated earlier
setwd("C:/Users/Ltischer/Documents/Studium/A Master/Geostatistics/Deforestation_SouthernGermany")
p193r27_1988_cdr_new_crop <- brick("results/p193r27_1988_cdr_new_crop.grd")
p193r27_1988_cdr_new_crop_illu <- brick("results/p193r27_1988_cdr_new_crop_illu.grd")


# plot to compare the results (2 plots in same window)
g1 <- ggRGB(p193r27_1988_cdr_new_crop, stretch="lin") +
            ggtitle("1988_crop\n(without illumination correction)")
g2 <- ggRGB(p193r27_1988_cdr_new_crop_illu, stretch="lin") +
            ggtitle("1988_crop\n(with illumination correction)")

grid.arrange(g1, g2, ncol=2, nrow=1)
```

```{r}
### check for clouds:


# process follows the device/description of package (https://cran.r-project.org/web/packages/RStoolbox/RStoolbox.pdf)
# only the cropped part of each scene should have been cloudMasked
# no big cloud contained on images --> finally no extra processing done.

########## 1. Data from 1988
cloudmask_1988_crop <- cloudMask(p193r27_1988_cdr_crop, blue=1, tir=6)
ggR(cloudmask_1988_crop, geom_raster=TRUE)
# --> no clouds identified

########## 2. Data from 2011
cloudmask_2011_crop <- cloudMask(p193r27_2011_cdr_crop, blue=1, tir=6)
ggR(cloudmask_2011_crop, geom_raster=TRUE)
# --> no clouds identified.
# --> no clouds removed.
```


***


### 5. Landcover Classification
  + Supervised Classification with RStoolbox::superClass()  
  + training and validation data collected in QGIS  
    (5 polygons per Landcover Class)
  + Landcover Classes:  
    - forest
    - water
    - urban/streets
    - agriculture/grassland
  + Raster data: cropped/masked to 25km-buffer around Marktoberdorf
  + Models: rf, svmRadial, pls (with and without NDVI)
  + Accuracy from validation-performance

```{r}
# Classification with supervised Classification (superClass RStoolbox).
# The classes 1: forest, 2: water, 3: urban/streets, 4: grassland/agriculture were distinguished.
# Training data consists of 5 polygons per class.
# Validation data also consists of 5 polygons per class.
# Classified cdr-raster data (Landsat 4-5 TM, Level 2) were illumination corrected ("C"-Method) 
# and masked to the 25km-buffer around Marktoberdorf.
# Chosen model was: "rf"

setwd("C:/Users/Ltischer/Documents/Studium/A Master/Geostatistics/Deforestation_SouthernGermany")

########## 1. read in the training- and validation data (defined in QGIS) and Raster scene
p193r27_1988_cdr_new_crop_illu_mask <- brick("results/p193r27_1988_cdr_new_crop_illu_mask.grd")
train_1988 <- readOGR(dsn="vector", layer="training_1988")
validate_1988 <- readOGR(dsn="vector", layer="validation_1988")

p193r27_2011_cdr_new_crop_illu_mask <- brick("results/p193r27_2011_cdr_new_crop_illu_mask.grd")
train_2011 <- readOGR(dsn="vector", layer="training_2011")
validate_2011 <- readOGR(dsn="vector", layer="validation_2011")



# make sure that the projections are identical
identical(projection(p193r27_1988_cdr_new_crop_illu_mask), projection(train_1988))
identical(projection(p193r27_1988_cdr_new_crop_illu_mask), projection(validate_1988))

identical(projection(p193r27_2011_cdr_new_crop_illu_mask), projection(train_2011))
identical(projection(p193r27_2011_cdr_new_crop_illu_mask), projection(validate_2011))




########## 2. Calucalte the NDVI per Scene and stack it to raster data
### NDVI 1988
p193r27_1988_cdr_NDVI <- spectralIndices(p193r27_1988_cdr_new_crop_illu_mask, 
                                         blue=1, red=2, green=3, nir=4, 
                                         indices="NDVI")
# stack NDVI with raster-bands
p193r27_1988_cdr_new_crop_illu_mask_NDVI <- stack(p193r27_1988_cdr_new_crop_illu_mask, p193r27_1988_cdr_NDVI)

# safe the "pure" NDVI values for later use
writeRaster(p193r27_1988_cdr_NDVI, "results/p193r27_1988_cdr_NDVI.grd")




### NDVI 2011
p193r27_2011_cdr_NDVI <- spectralIndices(p193r27_2011_cdr_new_crop_illu_mask, 
                                         blue=1, red=2, green=3, nir=4, 
                                         indices="NDVI")
# stack NDVI with raster-bands
p193r27_2011_cdr_new_crop_illu_mask_NDVI <- stack(p193r27_2011_cdr_new_crop_illu_mask, p193r27_2011_cdr_NDVI)
# safe the "pure" NDVI values for later use
writeRaster(p193r27_2011_cdr_NDVI, "results/p193r27_2011_cdr_NDVI.grd") 



# save the results
writeRaster(p193r27_1988_cdr_new_crop_illu_mask_NDVI, "results/p193r27_1988_cdr_new_crop_illu_mask_NDVI.grd")
writeRaster(p193r27_2011_cdr_new_crop_illu_mask_NDVI, "results/p193r27_2011_cdr_new_crop_illu_mask_NDVI.grd")

# reload results
p193r27_1988_cdr_new_crop_illu_mask_NDVI <- brick("results/p193r27_1988_cdr_new_crop_illu_mask_NDVI.grd")
p193r27_2011_cdr_new_crop_illu_mask_NDVI <- brick("results/p193r27_2011_cdr_new_crop_illu_mask_NDVI.grd")

p193r27_1988_cdr_NDVI <- raster("results/p193r27_1988_cdr_NDVI.grd")
p193r27_2011_cdr_NDVI <- raster("results/p193r27_2011_cdr_NDVI.grd")


########## 3. run the classifications

# classifications 1988
## !BEST! ##
set.seed(6)
p193r27_1988_cdr_sclass1 <- superClass(img = p193r27_1988_cdr_new_crop_illu_mask, 
                                       model = "rf", 
                                       trainData = train_1988, 
                                       valData = validate_1988,
                                       responseCol = "id")

set.seed(6)
p193r27_1988_cdr_sclass2 <- superClass(img = p193r27_1988_cdr_new_crop_illu_mask_NDVI, 
                                       model = "rf", 
                                       trainData = train_1988, 
                                       valData = validate_1988,
                                       responseCol = "id")

set.seed(6)
p193r27_1988_cdr_sclass3 <- superClass(img = p193r27_1988_cdr_NDVI, 
                                       model = "rf", 
                                       trainData = train_1988, 
                                       valData = validate_1988,
                                       responseCol = "id")

set.seed(6)
p193r27_1988_cdr_sclass4 <- superClass(img = p193r27_1988_cdr_new_crop_illu_mask, 
                                       model = "svmRadial", 
                                       trainData = train_1988, 
                                       valData = validate_1988,
                                       responseCol = "id")


set.seed(6)
p193r27_1988_cdr_sclass5 <- superClass(img = p193r27_1988_cdr_new_crop_illu_mask, 
                                       model = "pls", 
                                       trainData = train_1988, 
                                       valData = validate_1988,
                                       responseCol = "id")

set.seed(6)
p193r27_1988_cdr_sclass6 <- superClass(img = p193r27_1988_cdr_new_crop_illu_mask_NDVI, 
                                       model = "svmRadial", 
                                       trainData = train_1988, 
                                       valData = validate_1988,
                                       responseCol = "id")

p193r27_1988_cdr_sclass7 <- superClass(img = p193r27_1988_cdr_new_crop_illu_mask_NDVI, 
                                       model = "pls", 
                                       trainData = train_1988, 
                                       valData = validate_1988,
                                       responseCol = "id")




#classifications 2011
set.seed(6)
p193r27_2011_cdr_sclass1 <- superClass(img = p193r27_2011_cdr_new_crop_illu_mask, 
                                       model = "rf", 
                                       trainData = train_2011, 
                                       valData = validate_2011,
                                       responseCol = "id")

set.seed(6)
p193r27_2011_cdr_sclass2 <- superClass(img = p193r27_2011_cdr_new_crop_illu_mask_NDVI, 
                                       model = "rf", 
                                       trainData = train_2011, 
                                       valData = validate_2011,
                                       responseCol = "id")

set.seed(6)
p193r27_2011_cdr_sclass3 <- superClass(img = p193r27_2011_cdr_NDVI, 
                                       model = "rf", 
                                       trainData = train_2011, 
                                       valData = validate_2011,
                                       responseCol = "id")

set.seed(6)
p193r27_2011_cdr_sclass4 <- superClass(img = p193r27_2011_cdr_new_crop_illu_mask, 
                                       model = "svmRadial", 
                                       trainData = train_2011, 
                                       valData = validate_2011,
                                       responseCol = "id")

set.seed(6)
p193r27_2011_cdr_sclass5 <- superClass(img = p193r27_2011_cdr_new_crop_illu_mask, 
                                       model = "pls", 
                                       trainData = train_2011, 
                                       valData = validate_2011,
                                       responseCol = "id")

## !BEST! ##
set.seed(6)
p193r27_2011_cdr_sclass6 <- superClass(img = p193r27_2011_cdr_new_crop_illu_mask_NDVI, 
                                       model = "svmRadial", 
                                       trainData = train_2011, 
                                       valData = validate_2011,
                                       responseCol = "id")

set.seed(6)
p193r27_2011_cdr_sclass7 <- superClass(img = p193r27_2011_cdr_new_crop_illu_mask_NDVI, 
                                       model = "pls", 
                                       trainData = train_2011, 
                                       valData = validate_2011,
                                       responseCol = "id")





########## 4. Evaluate the quality of the Classifications

# accuracy for 1988
p193r27_1988_cdr_sclass1$validation$performance # Accuracy : 0.8162 ## rf without NDVI is best Model!
p193r27_1988_cdr_sclass1$modelFit
p193r27_1988_cdr_sclass2$validation$performance # Accuracy : 0.8113 # rf with NDVI
p193r27_1988_cdr_sclass3$validation$performance # Accuracy : 0.6237 # rf only with NDVI 
p193r27_1988_cdr_sclass4$validation$performance # Accuracy : 0.8043 # svmRadial 
p193r27_1988_cdr_sclass5$validation$performance # Accuracy : 0.7645 # pls
p193r27_1988_cdr_sclass6$validation$performance # Accuracy : 0.8067 # svmRaial with NDVI
p193r27_1988_cdr_sclass7$validation$performance # Accuracy : 0.7605 # pls with NDVI

# accuracy for 2011
p193r27_2011_cdr_sclass1$validation$performance # Accuracy : 0.8986 # rf 
p193r27_2011_cdr_sclass1$modelFit
p193r27_2011_cdr_sclass2$validation$performance # Accuracy : 0.8943 # rf with NDVI
p193r27_2011_cdr_sclass3$validation$performance # Accuracy : 0.6211 # rf only with NDVI
p193r27_2011_cdr_sclass4$validation$performance # Accuracy : 0.9155 # svmRadial
p193r27_2011_cdr_sclass5$validation$performance # Accuracy : 0.8355 # pls
p193r27_2011_cdr_sclass6$validation$performance # Accuracy : 0.9163 # svmRadial with NDVI is best Model
p193r27_2011_cdr_sclass7$validation$performance # Accuracy : 0.8348 # pls with NDVI


# save the maps of the best classifications
writeRaster(p193r27_1988_cdr_sclass1$map, "results/p193r27_1988_cdr_sclass1$map.grd")
writeRaster(p193r27_2011_cdr_sclass6$map, "results/p193r27_2011_cdr_sclass6$map.grd")
```


***


### Landcover Classification: Accuracies
|Year |NDVI     |rf        |svmRadial |pls   |
|:---:|:-------:|:--------:|:--------:|:----:|
|1988 |no       |**0.8162**|0.8043    |0.7645|
|1988 |included |0.8113    |0.8067    |0.7605|
|1988 |only     |0.6237    |-         |-     |
|2011 |no       |0.8986    |0.9155    |0.8355|
|2011 |included |0.8943    |**0.9163**|0.8348|
|2011 |only     |0.6211    |-         |-     |


***


### Landcover Classification: Result 1988

```{r, eval=TRUE}
setwd("C:/Users/Ltischer/Documents/Studium/A Master/Geostatistics/Deforestation_SouthernGermany")

# 2011: "new" lake in the southwest --> Rottachstausee --> planned and constructed in the 70ths and 80ths
# first time flooded: 1990
# regulary used since: 1992

# define the coordinates of Rottachsee
gps_Rottach <- as.data.frame(cbind( x=10.394951, y=47.652036, id=1))
coordinates(gps_Rottach) <- c("x", "y")
```

```{r}
projection(gps_Rottach)
projection(p193r27_1988_cdr)
```

```{r, eval=TRUE}
proj4string(gps_Rottach) <- CRS("+proj=longlat +datum=WGS84")
gps_Rottach <- spTransform(gps_Rottach, CRS(projection(p193r27_1988_cdr)))

```

```{r}
gps_Rottach
```

```{r, eval=TRUE, fig.height=6, fig.width=9.5}
setwd("C:/Users/Ltischer/Documents/Studium/A Master/Geostatistics/Deforestation_SouthernGermany")

# load results calculated earlier
p193r27_1988_cdr_sclass1_map <- raster("results/p193r27_1988_cdr_sclass1$map.grd")


# plot 1988 without the Rottachspeicher-Buffer
cols <- c("1"="seagreen3", "2"="blue", "3"="darkred", "4"="khaki1")

## 1988
ggR(p193r27_1988_cdr_sclass1_map, forceCat = TRUE, geom_raster = TRUE) +
  ggtitle(paste("Landscape 1988\nSupervised classification\n(model: rf)")) +
  theme(plot.title = element_text(size = 12, colour = "black", face="bold"), 
        legend.title= element_text(size=11, colour="black", face="bold")) +
  scale_fill_manual(values = cols, 
                    labels=c("Class1: Forest", "Class2: Water", "Class3: Urban/Streets", "Class4: Agriculture/Grassland"), 
                    name = "Landcover\nClasses\n")

```

***
### Landcover Classification: Result 2011

```{r, eval=TRUE, fig.height=6, fig.width=9.5}
setwd("C:/Users/Ltischer/Documents/Studium/A Master/Geostatistics/Deforestation_SouthernGermany")

# load results calculated earlier
p193r27_2011_cdr_sclass6_map <- raster("results/p193r27_2011_cdr_sclass6$map.grd")

#Create a 5km-Buffer around Rottachspeicher
buf_5 <- gBuffer(gps_Rottach, width=5000, byid=TRUE) 
#plot(buf_5)
#plot(gps_Rottach, add=TRUE)

# prepare the polygons for plotting
buf_5_df <- fortify(buf_5, region="id") # transform the buffer-data to dataframe
buf_5_df_final <- merge(buf_5_df, buf_5@data, by="id") # add the information

# transform also coordinates back to dataframe for plotting
gps_Rottach_df <- as.data.frame(gps_Rottach) 


## plot 2011 with buffer of Rottach-Speicher
cols <- c("1"="seagreen3", "2"="blue", "3"="darkred", "4"="khaki1")
ggR(p193r27_2011_cdr_sclass6_map, forceCat = TRUE, geom_raster = TRUE) +
  ggtitle(paste("Landscape 2011\nSupervised classification\n(model: svmRadial, with NDVI)")) +
  theme(plot.title = element_text(size = 12, colour = "black", face="bold"), 
        legend.title= element_text(size=11, colour="black", face="bold")) +
  scale_fill_manual(values = cols, 
                    labels=c("Class1: Forest", "Class2: Water", "Class3: Urban/Streets", "Class4: Agriculture/Grassland"), 
                    name = "Landcover\nClasses\n")+
  geom_polygon(data=buf_5_df_final, 
               mapping=aes(x=buf_5_df_final$long, 
                           y=buf_5_df_final$lat,
                           group=group), 
               colour="red", 
               fill="red", 
               alpha=0.1,
               lwd=2)

```


***


### Amount per Landcover Class

```{r, eval=TRUE}
# load results calculated earlier
setwd("C:/Users/Ltischer/Documents/Studium/A Master/Geostatistics/Deforestation_SouthernGermany")
p193r27_1988_cdr_new_crop_illu_mask <- brick("results/p193r27_1988_cdr_new_crop_illu_mask.grd")
p193r27_2011_cdr_new_crop_illu_mask <- brick("results/p193r27_2011_cdr_new_crop_illu_mask.grd")

########## 1. Area and Percentage of Landcover-Types in 1988
# count the pixels per class
p193r27_1988_cdr_sclass1.freq <- freq(p193r27_1988_cdr_sclass1_map, useNA="no")

# transform the Pixel-counts to surface-area
resLsat5_1988 <- res(p193r27_1988_cdr_new_crop_illu_mask) # query the spatial resolution of one pixel
area_km2_1988 <- p193r27_1988_cdr_sclass1.freq[ , "count"] * prod(resLsat5_1988) * 1e-6 # calculate the surface per class with: number of counts * pixelsize (30^2) * 1e-06 (in kilometers)
area_percent_1988 <- area_km2_1988/sum(area_km2_1988) # calculate also the %-Area per Lancover Type
area_km2_1988_df <- data.frame(Year=c(rep(1988,4)),
                               ClassID = c(1,2,3,4), 
                               Landcover = c("Forest", "Water", "Urban", "Agriculture"), 
                               area_km2 = area_km2_1988, 
                               area_percent = area_percent_1988)




########## 2. Area and Percetage of Landcover-Types in 2011
# count the pixels per class
p193r27_2011_cdr_sclass6.freq <- freq(p193r27_2011_cdr_sclass6_map, useNA="no")

# transform the Pixel-counts to surface-area
resLsat5_2011 <- res(p193r27_2011_cdr_new_crop_illu_mask) # query the spatial resolution of one pixel
area_km2_2011<- p193r27_2011_cdr_sclass6.freq[ , "count"] * prod(resLsat5_2011) * 1e-6 # calculate the surface per class with: number of counts * pixelsize (30^2) * 1e-06 (in kilometers)
area_percent_2011 <- area_km2_2011/sum(area_km2_2011) # calculate also the %-Area per Lancover Type
area_km2_2011_df <- data.frame(Year=c(rep(2011,4)),
                               ClassID = c(1,2,3,4), 
                               Landcover = c("Forest", "Water", "Urban", "Agriculture"), 
                               area_km2 = area_km2_2011, 
                               area_percent = area_percent_2011)

# connect the two dataframes to one
area_km2_df_complete <- rbind(area_km2_1988_df, area_km2_2011_df)





########## 3. Plotting the Percentage of Landcover Types

cols_2 <- c("Forest"="seagreen3", "Water"="blue", "Urban"="brown1", "Agriculture"="khaki1")

ggplot(data=area_km2_df_complete, 
       aes(x=as.factor(Year), 
           y=area_percent*100, 
           fill=Landcover, 
           label=round(area_percent*100, 2))) +
  geom_bar(stat="identity", 
           width=0.7) + 
  #geom_text(size = 4.5, 
  #          position = position_stack(vjust = 1.1), 
  #          colour="black") +
  geom_text(size = 4, hjust = 0.5, vjust = 1.3, position = "stack") +
  scale_fill_manual(values =cols_2) +
  ggtitle("Percentage of Landcover Classes\n(1988 vs. 2011)\n") +
  xlab("\n\nYear") + 
  ylab("Amount per Landcover Class [%]\n\n") +
  theme(plot.title = element_text(size = 15, colour = "black", face="bold"), 
        legend.title=element_text(size = 14, colour = "black", face="bold"),
        legend.text = element_text(size = 12, colour = "black", face="plain"),
        axis.title.y = element_text(size = 14, colour = "black", face="bold"), 
        axis.title.x = element_text(size = 14, colour = "black", face="bold"), 
        axis.text.x = element_text(size = 13, colour = "black", face="bold"), 
        axis.text.y = element_text(size = 13, colour = "black", face="bold"))
```


***

### Productivity per Landcover Class
```{r, eval=TRUE}
setwd("C:/Users/Ltischer/Documents/Studium/A Master/Geostatistics/Deforestation_SouthernGermany")

# load results calculated earlier
p193r27_1988_cdr_sclass1_map <- raster("results/p193r27_1988_cdr_sclass1$map.grd")
p193r27_2011_cdr_sclass6_map <- raster("results/p193r27_2011_cdr_sclass6$map.grd")

# load the NDVI-Values
p193r27_1988_cdr_NDVI <- raster("results/p193r27_1988_cdr_NDVI.grd")
p193r27_2011_cdr_NDVI <- raster("results/p193r27_2011_cdr_NDVI.grd")

# group the single pixel- NDVI-values per class
forest_prod_1988 <- p193r27_1988_cdr_NDVI[p193r27_1988_cdr_sclass1_map$id == 1,]
water_prod_1988 <- p193r27_1988_cdr_NDVI[p193r27_1988_cdr_sclass1_map$id == 2,]
urban_prod_1988 <- p193r27_1988_cdr_NDVI[p193r27_1988_cdr_sclass1_map$id == 3,]
agriculture_prod_1988 <- p193r27_1988_cdr_NDVI[p193r27_1988_cdr_sclass1_map$id == 4,]
prod_1988 <- cbind(forest=forest_prod_1988, 
                   water=water_prod_1988, 
                   urban=urban_prod_1988, 
                   agriculture=agriculture_prod_1988)


forest_prod_2011 <- p193r27_2011_cdr_NDVI[p193r27_2011_cdr_sclass6_map$id == 1, ]
water_prod_2011 <- p193r27_2011_cdr_NDVI[p193r27_2011_cdr_sclass6_map$id == 2, ]
urban_prod_2011 <- p193r27_2011_cdr_NDVI[p193r27_2011_cdr_sclass6_map$id == 3, ]
agriculture_prod_2011 <- p193r27_2011_cdr_NDVI[p193r27_2011_cdr_sclass6_map$id == 4, ]
prod_2011 <- cbind(forest=forest_prod_2011, 
                   water=water_prod_2011, 
                   urban=urban_prod_2011, 
                   agriculture=agriculture_prod_2011)


# plot the Productivity per class and year
par(mfrow=c(1,2), mar=c(6,5,4,2), lwd=1, oma=c(0,0,4,0))
boxplot(prod_1988,
        ylim=c(-1.1, 1.1),
        ylab="", 
        xlab="", 
        names=c("Forest", "Water", "Urban", "Agriculture"), 
        las=2, 
        cex.axis=1.0,
        col = c("seagreen3", "blue", "darkred", "khaki1"))
mtext( "Landcover Classes",
       side = 1,line = 5,cex = 1.1, las = 0, font = 2)
mtext( "NDVI-Values",
       side = 2,line = 3,cex = 1.1, las = 0, font = 2)
mtext("Comparison of Productivity\n(1988)",
      side = 3,line = 1, cex = 1.2, font = 2)


boxplot(prod_2011,
        ylim=c(-1.1, 1.1), 
        ylab="", 
        xlab="", 
        names=c("Forest", "Water", "Urban", "Agriculture"), 
        las=2, 
        cex.axis=1.0,
        col = c("seagreen3", "blue", "darkred", "khaki1"))
mtext( "Landcover Classes",
       side = 1,line = 5,cex = 1.1, las = 0, font = 2)
mtext( "NDVI-Values",
       side = 2,line = 3,cex = 1.1, las = 0, font = 2)
mtext("Comparison of Productivity\n(2011)",
      side = 3,line = 1, cex = 1.2, font = 2)
mtext("Comparison between 1988 and 2011",
      side = 3,line = 1, cex = 1.2, font = 2, outer=TRUE)


# reset the plotting parameters to default
#dev.off()
```


***

### Significance? (Productivity ~ Landcover Class, Year)

```{r}
# test for Normal distribution
# therefore: take 5000 samples from all values (otherwise to much data for shapiro.test)
set.seed(6)
forest_prod_1988_sample <- sample(forest_prod_1988, 5000, replace=FALSE) 
set.seed(6)
agriculture_prod_1988_sample <- sample(agriculture_prod_1988, 5000, replace=FALSE) 
set.seed(6)
forest_prod_2011_sample <- sample(forest_prod_2011, 5000, replace=FALSE) 
set.seed(6)
agriculture_prod_2011_sample <- sample(agriculture_prod_2011, 5000, replace=FALSE) 

# run the actual tests
shapiro.test(forest_prod_1988_sample) # W = 0.94261, p-value < 2.2e-16      
shapiro.test(agriculture_prod_1988_sample) # W = 0.97013, p-value < 2.2e-16
shapiro.test(forest_prod_2011_sample) # W = 0.94333, p-value < 2.2e-16
shapiro.test(agriculture_prod_2011_sample) # W = 0.93058, p-value < 2.2e-16
# --> none is normal distributed!!

# check for Homogenity of Variance
bartlett.test(list(forest_prod_1988_sample, agriculture_prod_1988_sample)) # Bartlett's K-squared = 922.66, df = 1, p-value < 2.2e-16
bartlett.test(list(forest_prod_1988_sample, forest_prod_2011_sample)) # Bartlett's K-squared = 181.8, df = 1, p-value < 2.2e-16
bartlett.test(list(agriculture_prod_1988_sample, agriculture_prod_2011_sample)) # Bartlett's K-squared = 72.818, df = 1, p-value < 2.2e-16
bartlett.test(list(agriculture_prod_2011_sample, forest_prod_2011_sample)) # Bartlett's K-squared = 654.54, df = 1, p-value < 2.2e-16
# --> none is with homogenous variance!!


# alternative for t-test for non-parametric data: Mann-Whitney-U Test
##### 1. test, H0: they are the same. H1: forest_1988 is higher than forest_2011
wilcox.test(forest_prod_1988_sample, forest_prod_2011_sample, alternative="greater") 
# W = 13927000, p-value < 2.2e-16 --> forest_1988 is higher!

##### 2. H1: forst_1988 is higher than agriculture 1988
wilcox.test(forest_prod_1988_sample, agriculture_prod_1988_sample, alternative="greater") 
# W = 13212000, p-value = 7.983e-07  --> forest_1988 is higher!

##### 3. H1: forest_2011 is higher than agriculture_2011
wilcox.test(forest_prod_2011_sample, agriculture_prod_2011_sample, alternative="greater") 
# W = 14656000, p-value < 2.2e-16  --> forest_2011 is higher!

##### 4. H1: agriculture_1988 is higher than agriculture_2011
wilcox.test(agriculture_prod_1988_sample, agriculture_prod_2011_sample, alternative="greater") 
# W = 14824000, p-value < 2.2e-16  --> forest_1988 is higher!
```
  
  
  + test for normal distribution: shapiro.test()
  + test for homegenous variance: bartlett.test()

**data not suited for t-test...**

  + Comparison between groups with **wilcox.test()**  
    (5000 random samples per group: requirement for test)  
    - forest 1988 > forest 2011:            p < 0.001  
    - forest 1988 > agriculture 1988:       p < 0.001  
    - forest 2011 > agriculture 2011:       p < 0.001  
    - agriculture 1988 > agriculture 2011:  p < 0.001
  

***

### 6. Elevation Zones: Topografic Profile

```{r, eval=TRUE}

########## 1. Loading the Data and Plotting
setwd("C:/Users/Ltischer/Documents/Studium/A Master/Geostatistics/Deforestation_SouthernGermany")

# load elevation and Raster-data
elev_germany_proj_crop_resampled_mask <- raster("results/elev_germany_proj_crop_resampled_mask.grd")
elev_germany_proj_crop_resampled_mask_df <- as.data.frame(elev_germany_proj_crop_resampled_mask)

p193r27_1988_cdr_new_crop_illu_mask <- brick("results/p193r27_1988_cdr_new_crop_illu_mask.grd")
p193r27_2011_cdr_new_crop_illu_mask <- brick("results/p193r27_2011_cdr_new_crop_illu_mask.grd")

p193r27_1988_cdr_NDVI <- raster("results/p193r27_1988_cdr_NDVI.grd")
p193r27_2011_cdr_NDVI <- raster("results/p193r27_2011_cdr_NDVI.grd")


# plot the elevation data as raster and as boxplot
## raster
ggR(elev_germany_proj_crop_resampled_mask, geom_raster = TRUE) +
  ggtitle("Elevation Data - Raster Image") +
  theme(plot.title = element_text(size = 13, colour = "black", face="bold"), 
        legend.title=element_text(size = 11, colour = "black", face="bold")) +
  scale_fill_gradientn(colours=terrain.colors(50), 
                       name = "Elevation [m]\n")
```


***

### Elevation Zones: Data Distribution
```{r, eval=TRUE, fig.height=4.5}
## boxplot   
par(mar=c(3,6,4,3))
boxplot(elev_germany_proj_crop_resampled_mask_df, 
        main="", 
        ylab="", 
        las=2, 
        cex.axes=1.1, 
        cex.lab=1.2, 
        col="azure2") 
mtext( "Elevation Values - Data Range",
       side = 3,line = 2,cex = 1.2, las = 0, font = 2)
mtext( "Elevation [m]",
       side = 2,line = 4,cex = 1.1, las = 0, font = 2)

# --> 2 big groups of elevation data can be distinguished: > 800m  and < 800m
# --> <= 800m productivity will be compared with > 800m productivity
```

  - big part of data between 600 m and 1000 m elevation
  - two groups chosen: <= 800 m and elevation > 800 m  


***

### Elevation Zones: Productivity (Forest)

```{r, eval=TRUE}

# hereafter: calculation of both forest and agriculture. plotting for forest.
# plotting agriculture: next chunk.

######### 2. Defining the Elevation Classes <= 800m and > 800m

low_elevation <- elev_germany_proj_crop_resampled_mask[elev_germany_proj_crop_resampled_mask <= 800]
high_elevation <- elev_germany_proj_crop_resampled_mask[elev_germany_proj_crop_resampled_mask > 800]
#length(low_elevation)
#length(high_elevation)




########## 3. define the productivity (agricultural area) per elevation class and year

# group the data depending on landcover class (forest vs. agriculture) and elevation hight per year
### 1988
forest_prod_1988_low <- p193r27_1988_cdr_NDVI[(p193r27_1988_cdr_sclass1_map$id == 1) & (elev_germany_proj_crop_resampled_mask <= 800),]
forest_prod_1988_high <- p193r27_1988_cdr_NDVI[(p193r27_1988_cdr_sclass1_map$id == 1) & (elev_germany_proj_crop_resampled_mask > 800),]
agriculture_prod_1988_low <- p193r27_1988_cdr_NDVI[(p193r27_1988_cdr_sclass1_map$id == 4) & (elev_germany_proj_crop_resampled_mask <= 800),]
agriculture_prod_1988_high <- p193r27_1988_cdr_NDVI[(p193r27_1988_cdr_sclass1_map$id == 4) & (elev_germany_proj_crop_resampled_mask > 800),]

prod_1988_forest_low_high <- cbind(forest_low=forest_prod_1988_low,
                                   forest_high=forest_prod_1988_high)
prod_1988_agriculture_low_high <- cbind(agriculture_low=agriculture_prod_1988_low, 
                                        agriculture_high=agriculture_prod_1988_high)


### 2011
forest_prod_2011_low <- p193r27_2011_cdr_NDVI[(p193r27_2011_cdr_sclass6_map$id == 1) & (elev_germany_proj_crop_resampled_mask <= 800),]
forest_prod_2011_high <- p193r27_2011_cdr_NDVI[(p193r27_2011_cdr_sclass6_map$id == 1) & (elev_germany_proj_crop_resampled_mask > 800),]
agriculture_prod_2011_low <- p193r27_2011_cdr_NDVI[(p193r27_2011_cdr_sclass6_map$id == 4) & (elev_germany_proj_crop_resampled_mask <= 800),]
agriculture_prod_2011_high <- p193r27_2011_cdr_NDVI[(p193r27_2011_cdr_sclass6_map$id == 4) & (elev_germany_proj_crop_resampled_mask > 800),]

prod_2011_forest_low_high <- cbind(forest_low=forest_prod_2011_low,
                                   forest_high=forest_prod_2011_high)
prod_2011_agriculture_low_high <- cbind(agriculture_low=agriculture_prod_2011_low, 
                                        agriculture_high=agriculture_prod_2011_high)


# plot everything
par(mfrow=c(1,2), mar=c(2,2,2,2), oma=c(0,4,6,0), las=1, cex.axis=1.2)
p1 <- boxplot(prod_1988_forest_low_high , ylim=c(0, 1), names=c("<= 800m", ">800 m"))
lines(x=c(0,1), y=c(p1$stats[3,1], p1$stats[3,1]), 
      col="red", 
      lty="solid", 
      lwd=2)
lines(x=c(0,2), y=c(p1$stats[3,2], p1$stats[3,2]), 
      col="gold", 
      lty="dashed", 
      lwd=2)
text(x=2.2, y=0.05,labels="Forest\n1988", cex=1.1)

p2 <- boxplot(prod_2011_forest_low_high, ylim=c(0, 1), names=c("<= 800m", ">800 m"))
lines(x=c(0,1), y=c(p2$stats[3,1], p2$stats[3,1]), 
      col="red", 
      lty="solid", 
      lwd=2)
lines(x=c(0,2), y=c(p2$stats[3,2], p2$stats[3,2]), 
      col="gold", 
      lty="dashed", 
      lwd=2)
text(x=2.2, y=0.05,labels="Forest\n2011", cex=1.1)
mtext( "NDVI-Values",
       side = 2,line = 2,cex = 1.1, las = 0, font = 2, outer =TRUE)
mtext("Productivity of forest\nin 1988 and 2011\n(per elevation class)",
      side = 3,line = 0, cex = 1.2, font = 2, outer=TRUE)

```

***

### Elevation Zones: Productivity (Agriculture)
```{r, eval=TRUE}
par(mfrow=c(1,2), mar=c(2,2,2,2), oma=c(0,4,6,0), las=1, cex.axis=1.2)
p3 <- boxplot(prod_1988_agriculture_low_high, ylim=c(0, 1), names=c("<= 800m", ">800 m"))
lines(x=c(0,1), y=c(p3$stats[3,1], p3$stats[3,1]), 
      col="red", 
      lty="solid", 
      lwd=2)
lines(x=c(0,2), y=c(p3$stats[3,2], p3$stats[3,2]), 
      col="gold", 
      lty="dashed", 
      lwd=2)
text(x=2.2, y=0.05,labels="Agriculture\n1988", cex=1.1)


p4 <- boxplot(prod_2011_agriculture_low_high, ylim=c(0, 1), names=c("<= 800m", ">800 m"))
lines(x=c(0,1), y=c(p4$stats[3,1], p4$stats[3,1]), 
      col="red", 
      lty="solid", 
      lwd=2)
lines(x=c(0,2), y=c(p4$stats[3,2], p4$stats[3,2]), 
      col="gold", 
      lty="dashed", 
      lwd=2)
text(x=2.2, y=0.05,labels="Agriculture\n2011", cex=1.1)
mtext( "NDVI-Values",
       side = 2,line = 2,cex = 1.1, las = 0, font = 2, outer =TRUE)
mtext("Productivity of agriculture\nin 1988 and 2011\n(per elevation class)",
      side = 3,line = 0, cex = 1.2, font = 2, outer=TRUE)
```

***

### Significance? (Productivity ~ Elevation, for both years)

```{r}
# hereafter: calculation Significances forest AND agriculture

### 1988
# take 5000 samples per group (otherwise too much for shapiro.test())
forest_prod_1988_low_sample <- sample(forest_prod_1988_low, 5000, replace=FALSE)
forest_prod_1988_high_sample <- sample(forest_prod_1988_high, 5000, replace=FALSE)
agriculture_prod_1988_low_sample <- sample(agriculture_prod_1988_low, 5000, replace=FALSE)
agriculture_prod_1988_high_sample <- sample(agriculture_prod_1988_high, 5000, replace=FALSE)

# test for normal distribution
shapiro.test(forest_prod_1988_low_sample) # W = 0.93993, p-value < 2.2e-16
shapiro.test(forest_prod_1988_high_sample) # W = 0.98321, p-value < 2.2e-16
shapiro.test(agriculture_prod_1988_low_sample) # W = 0.96364, p-value < 2.2e-16
shapiro.test(agriculture_prod_1988_high_sample) # W = 0.97147, p-value < 2.2e-16
# --> none is normal distributed  --> not suited for t-test

# Comparison of groups again with Mann-Witney-U Test
##### 1. H0: both are the same. H1: forest_low is higher than forest_high
wilcox.test(forest_prod_1988_low_sample, forest_prod_1988_high_sample, alternative="greater")
# W = 13870000, p-value < 2.2e-16

##### 2. H1: agriculture_low is higher than agriculture_high
wilcox.test(agriculture_prod_1988_low_sample, agriculture_prod_1988_high_sample, alternative="greater")
# W = 12731000, p-value = 0.05494




### 2011
# take 5000 samples per group (otherwise too much for shapiro.test())
forest_prod_2011_low_sample <- sample(forest_prod_2011_low, 5000, replace=FALSE)
forest_prod_2011_high_sample <- sample(forest_prod_2011_high, 5000, replace=FALSE)
agriculture_prod_2011_low_sample <- sample(agriculture_prod_2011_low, 5000, replace=FALSE)
agriculture_prod_2011_high_sample <- sample(agriculture_prod_2011_high, 5000, replace=FALSE)


# test for normal distribution
shapiro.test(forest_prod_2011_low_sample) # W = 0.95701, p-value < 2.2e-16
shapiro.test(forest_prod_2011_high_sample) # W = 0.96761, p-value < 2.2e-16
shapiro.test(agriculture_prod_2011_low_sample) # W = 0.89, p-value < 2.2e-16
shapiro.test(agriculture_prod_2011_high_sample) # W = 0.9676, p-value < 2.2e-16
# --> none is normal distributed  --> not suited for t-test


# Comparison of groups again with Mann-Witney-U Test
##### 1. H0: both are the same. H1: forest_low is higher than forest_high
wilcox.test(forest_prod_2011_low_sample, forest_prod_2011_high_sample, alternative="greater")
# W = 13097000, p-value = 1.757e-05

##### 2. H1: agriculture_low is higher than agriculture_high
wilcox.test(agriculture_prod_2011_low_sample, agriculture_prod_2011_high_sample, alternative="greater")
# W = 12637000, p-value = 0.1712

```

  - test for normal distribution: shapiro.test()

**data not suited for t-test...**

  + Comparison between groups with **wilcox.test()**  
    (5000 random samples per group: requirement for test)
  
  + **1988:**
    - forest 1988 low > forest 1988 high:       p < 0.001
    - agricul. 1988 low > agricult. 1988 high:  p = 0.0549
    
  + **2011:**
    - forest 2011 low > forest 2011 high:       p < 0.001
    - agricul. 2011 low > agricult. 2011 high:  p = 0.171


***

### 7. Summary
  
    
  - only little changes in landscape composition
    
  - mostly agriculture changed to urban  
    (growing cities)
      
  - little but significant difference between productivity  
    (~ landcover class, year and elevation)
    + forest > agriculture
    + 1988 > 2011
    + low forest > high forest (both years)
    + low agriculture = high agriculture (both years)  


***

### 8. Discussion
  - Satellite data of July
    + most productive season of the year
    + differences of prod maybe more pronounced in spring or autumn  
      (higher elevation still/already quite cold,   
       agricultural fields not yet fully covered)
  - grassland and agriculture as one class
    + grassland covers big parts of that region
    + grassland probably mostly unchanged since 1988  
      (**BUT:** grassland and agriculture hard to differentiate on Raster Image, 
                no GoogleSatellite overlay for old data)  
  - only one image of one region of one timestap compared  
    + not enough data for generall assumptions
              
***

### Outlook
  - Improvements of classifications concerning the date and the separation in landcover classes
  
  - Similar analysis in different regions/ years. Same Trends?

***

### End



**Thank** **you** **for** **your** **attention!!**
 



My GitHub: https://github.com/LuciaTi